---
title: "GDM"
author: ""
date: "Wednesday, August 05, 2015"
output: html_document
---


```{r, echo = FALSE, include = FALSE}
  
  load("C:/Users/ddahly/Dropbox/smallProjects/cumhDietetics/data.rda")
  load("C:/Users/ddahly/Dropbox/smallProjects/cumhDietetics/map.rda")


	library(tidyr)
	library(plyr)
	library(dplyr)
	library(ggplot2)
	library(RColorBrewer)
	library(readxl)
	library(gmodels)
	library(magrittr)
	library(xtable)
	library(rworldmap)

```

The data are on 406 women with singleton pregnancies. 

##GDM diagnosis##

GA at time of GTT

```{r, echo = FALSE, warning = FALSE}

  ggplot(data, aes(x = GAatGTT)) +
  	geom_bar(fill = brewer.pal(3, "Set1")[1]) +
  	ggtitle(paste("Gestational age when GTT was given (n = ",
  								nrow(data[!is.na(data$GAatGTT), ]),
  								")", sep = "")) +
  	ylab("Number of patients") +
  	xlab("Gestational age (weeks)")


```

GTT location

```{r, echo = FALSE, warning = FALSE}

  table(data$GTTLocation, dnn = "Location of GTT")

```

Distribution of fasting glucose (mmol)

```{r, echo = FALSE, warning = FALSE}
# GMD diagnosis ################################################################

# fasting of 5.1 or greater or  2 hour value of 8.5mmol or above.

  summary(data$Fasting)

```
  
Distribution of 2-hour glucose (mmol)

```{r, echo = FALSE, warning = FALSE}
# GMD diagnosis ################################################################

# fasting of 5.1 or greater or  2 hour value of 8.5mmol or above.

  summary(data$TwoHRPP)
```

Distribution of 1-hour glucose (mmol)

```{r, echo = FALSE, warning = FALSE}
# GMD diagnosis ################################################################

# fasting of 5.1 or greater or  2 hour value of 8.5mmol or above.

  summary(data$OneHRPP)
```

This is a plot of fasting and 2-hour gluose values, for women with both measures. 
The blue markers are women with 1-hour glucose >= 10 mmol. There are a few
women who don't seem to meet any of these criteria. 

```{r, echo = FALSE, warning = FALSE}

  ggplot(data, aes(x = Fasting, y = TwoHRPP)) +
  	geom_point(alpha = 0.7, size = 4, color = brewer.pal(3, "Set1")[1]) +
  	geom_point(data = filter(data, OneHRPP >= 10), size = 4,
  						 color = brewer.pal(3, "Set1")[2]) +
  	geom_hline(yintercept = 8.5) +
  	geom_vline(xintercept = 5.1) +
  	ggtitle(paste("Fasting and 2-hour blood glucose (n = ",
  								nrow(data[!is.na(data$Fasting) & !is.na(data$TwoHRPP), ]),
  								")", sep = "")) +
  	xlab("Fasting (mmol)") +
  	ylab("2-hour (mmol)")
```

Who are the women that don't meet the fasting or 2-hour gluose criteria?

```{r, echo = FALSE, warning = FALSE}

  filter(data[, c(1, 56:59)], TwoHRPP < 8.5 & Fasting < 5.1)

```

Distribution of HBA1c

```{r, echo = FALSE, warning = FALSE}
  summary(data$HBA1catdiagnosis)

  ggplot(data, aes(x = HBA1catdiagnosis)) +
  	geom_bar(fill = brewer.pal(3, "Set1")[1]) +
  	ggtitle(paste("HBA1c (n = ",
  								nrow(data[!is.na(data$HBA1c), ]),
  								")", sep = "")) +
  	ylab("Number of patients") +
  	xlab("HBA1c")

```


##Management##


Insulin Dose at end 

```{r, echo = FALSE, warning = FALSE}

 table(data$DietOnly, dnn = "Diet only")

```


```{r, echo = FALSE, warning = FALSE}

 table(data$ReferraltoEndocrinology, dnn = "Referral to Endocrinology")


```


```{r, echo = FALSE, warning = FALSE}

 table(data$Metformin, dnn = "Metformin")

```


```{r, echo = FALSE, warning = FALSE}

 table(data$BasalInsulin, dnn = "Basal Insulin")


```


```{r, echo = FALSE, warning = FALSE}

 table(data$PrandialInsulin, dnn = "Prandial Insulin")


```

Compare all of these. 

Metformin vs Insulin

```{r, echo = FALSE, warning = FALSE}

 table(data$PrandialInsulin, data$Metformin, 
 			dnn = c("Prandial Insulin", "Metformin"))
 			
 table(data$BasalInsulin, data$Metformin, 
 			dnn = c("Basal Insulin", "Metformin"))
 			
 table(data$PrandialInsulin, data$BasalInsulin, 
 			dnn = c("Prandial Insulin", "Basal Insulin")) 			

```


Diet vs metformin/insulin

```{r, echo = FALSE, warning = FALSE}

 table(data$DietOnly, data$Metformin, 
 			dnn = c("Diet", "Metformin"))
 			
 table(data$DietOnly, data$PrandialInsulin, 
 			dnn = c("Diet", "Prandial Insulin"))
 			
 table(data$DietOnly, data$BasalInsulin, 
 			dnn = c("Diet", "Basal Insulin")) 			

```




Insulin dose at the end: total sample, in those with Basal Insulin, and in those with Prandial Insulin. 

```{r, echo = FALSE, warning = FALSE}

  summary(data$InsulinDoseatend)

  summary(data[data$BasalInsulin == "Yes", ]$InsulinDoseatend)

	ggplot(data[data$BasalInsulin == "Yes", ], aes(x = InsulinDoseatend)) +
		geom_bar(fill = brewer.pal(3, "Set1")[1]) +
		ggtitle(paste("HBA1c (n = ",
									nrow(data[data$BasalInsulin == "Yes" & 
															!is.na(data$InsulinDoseatend), ]),
									")", sep = "")) +
		ylab("Number of patients") +
		xlab("Insulin dose")


  summary(data[data$PrandialInsulin == "Yes", ]$InsulinDoseatend)

	ggplot(data[data$PrandialInsulin == "Yes", ], aes(x = InsulinDoseatend)) +
		geom_bar(fill = brewer.pal(3, "Set1")[1]) +
		ggtitle(paste("HBA1c (n = ",
									nrow(data[data$PrandialInsulin == "Yes" & 
															!is.na(data$InsulinDoseatend), ]),
									")", sep = "")) +
		ylab("Number of patients") +
		xlab("Insulin dose")


```








##GDM screening/risk factors##

Previous history of GDM. This is based on the variable PHxGDM. Women with parity
0 are labelled firstborn. 

```{r, echo = FALSE, warning = FALSE}

  table(data$prevGDM,	dnn = c("Previous history of GMD")) 

```

Other factors for screening that we have data on are:

* Family history of diabetes in a first degree relative
* Body mass index ≥30kg/m2
* Maternal age ≥ 40years
* Previous unexplained perinatal death
* Current glycosuria
* Women on long term steroids
* Previous delivery of a baby weighing ≥4.5kg
* Polycystic Ovary Syndrome
* Polyhydramnios and/or macrosomia in existing pregnancy
* Ethnicity associated with a high prevalence of diabetes


###Family history of diabetes in a first degree relative###

```{r, echo = FALSE, warning = FALSE}

	table(data$FHxT2DM, data$FHxT1DM,
  			dnn = c("FHx Type II Diabetes", "FHx Type I Diabetes")) 

  table(data$anyFHxDm, dnn = "Any family history of diabetes")

```

###Heights and weights (used to calculate BMI)###

There were a few data entry errors that have been corrected. Here is a plot of
the heights and weights. 

```{r, echo = FALSE, warning = FALSE}

	ggplot(data, aes(x = Height, y = Weight)) +
		geom_jitter(alpha = 0.7, size = 4, color = brewer.pal(3, "Set1")[1]) +
  	ggtitle(paste("Heights and weights (n = ",
  								nrow(data[!is.na(data$Height) & !is.na(data$Weight), ]),
                  ")", sep = "")) +
  	xlab("Height (m)") +
  	ylab("Weight (kg)")

```

###BMI###

Based on the corrected heights and weights, I calculated a new BMI variable. It
does not match up perfectly with the BMI values in the dataset. Here is a plot
to illustrate this. 

```{r, echo = FALSE, warning = FALSE}

 ggplot(data, aes(x = BMI, y = BMInew)) +
  	geom_jitter(alpha = 0.7, size = 4, color = brewer.pal(3, "Set1")[1]) +
  	ggtitle(paste("Original and new BMI values (n = ",
  								nrow(data[!is.na(data$BMI) & !is.na(data$BMInew), ]),
  								")", sep = "")) +
  	xlab("Original BMI value (kg/m2)") +
  	ylab("New BMI value (kg/m2)")

```

Some of the differences are large. 

```{r, echo = FALSE, warning = FALSE}

 ggplot(data, aes(x = bmiDifference)) +
  	geom_bar(binwidth = 0.1, fill = brewer.pal(3, "Set1")[1]) +
  	xlab("Difference between BMI values")


```

Because of these differences, I have also created a new body size classification variable. Here is how it matches up with the original version in the dataset. 

```{r, echo = FALSE, warning = FALSE}

	table(data$BMIclass, data$BMIclassNew,
  			dnn = c("Original BMI", "Newly calculated BMI"))

```

###Maternal age###

```{r, echo = FALSE, warning = FALSE}
  
	summary(data$Age)

  ggplot(data, aes(x = Age)) +
  	geom_bar(fill = brewer.pal(3, "Set1")[1]) +
  	ggtitle(paste("Maternal Ages (n = ",
  								nrow(data[!is.na(data$Age), ]),
  								")", sep = "")) +
  	ylab("Number of patients") +
  	xlab("Age (years)")

  table(data$fourty, dnn = "Maternal age >= 40 years")

```

###Weight of the previous baby###

I created a new variable which reflected a macrosomic previous baby if the variable LargeBaby indicated so, or if the variable Weightofbiggestbaby was >= 4500g. Women with a parity of zero are labelled as firstborn. 

Here is a plot of the weights of the biggest previous babies. 

```{r, echo = FALSE, warning = FALSE}

 	ggplot(data, aes(x = Weightofbiggestbaby)) +
		geom_bar(fill = brewer.pal(3, "Set1")[1]) +
		ggtitle(paste("Birth weight of largest previous baby (n = ",
									nrow(data[!is.na(data$Weightofbiggestbaby), ]),
	                ")", sep = "")) +
		ylab("Number of patients") +
		xlab("Birth weight (g)")
 

```

A table of LargeBaby and Parity

```{r, echo = FALSE, warning = FALSE}

 table(data$LargeBaby, data$Parity, dnn = c("Large baby", "Parity")) 


```

The new variable - not much difference. 

```{r, echo = FALSE, warning = FALSE}

  table(data$bigBaby, dnn = "Previous macrocomic birth")

```

Here is the distribution of Weightofbiggestbaby by LargeBaby

```{r, echo = FALSE, warning = FALSE}

   data[ , !duplicated(colnames(data))] %>%
  	group_by(as.factor(LargeBaby)) %>%
  	summarise(n    = n(),
  		        mean = mean(Weightofbiggestbaby, na.rm = TRUE),
  						min =  min(Weightofbiggestbaby,  na.rm = TRUE),
  						max =  max(Weightofbiggestbaby,  na.rm = TRUE))

```


###PCOS###

```{r, echo = FALSE, warning = FALSE}
 
  table(data$PHxPCOS, dnn = "History of PCOS")

```


###Polyhydraminos###

```{r, echo = FALSE, warning = FALSE}

  table(data$polyhydraminos, dnn = "Polyhydraminos")

```


###Curent Macro###
Any AC reading >= 90th centile for GA

```{r, echo = FALSE, warning = FALSE}

 table(data$suspectMacro, dnn = "Suspected macrosomia")

```

###Ethicity###

Based on black ethnicity, or country of origin being in Africa (expect South Africa), or on the list in the screening guidelines.

```{r, echo = FALSE, warning = FALSE}

  table(data$ethnicRisk, dnn = "Risky ethnicity")

```



##Socio/Demographics/Health##

###County and distance to CUMH###

```{r, echo = FALSE, warning = FALSE}

	table(data$County, dnn = "County of residence")

  ggplot(data, aes(x = County, fill = County)) +
  	geom_bar() +
  	scale_fill_brewer("", palette = "Set1", guide = FALSE) +
  	ggtitle(paste("County of residence (n = ",
  								nrow(data[!is.na(data$County), ]), ")", sep = "")) +
  	xlab("County") +
  	ylab("Number of participants")

# Distance ####

  summary(data$Distance)

  ggplot(data, aes(x = DistancefromCUMHKm)) +
  	geom_bar(fill = brewer.pal(3, "Set1")[1]) +
  	ggtitle(paste("Distance from CUMH (n = ",
  								nrow(data[!is.na(data$DistancefromCUMHKm), ]),
  								")", sep = "")) +
  	xlab("Distance (km)") +
  	ylab("Number of participants")

```

###Ethnicity###

```{r, echo = FALSE, warning = FALSE}

  table(data$RacialCategory, dnn = "Ethnicity")
  
  ggplot(data, aes(x = RacialCategory, fill = RacialCategory)) +
  	geom_bar() +
  	scale_fill_brewer("", palette = "Set1", guide = FALSE) +
  	ggtitle(paste("Ethnicity (n = ",
  								nrow(data[!is.na(data$RacialCategory), ]), ")", sep = "")) +
  	xlab("Ethnicity") +
  	ylab("Number of participants")

  table(data$white, data$RacialCategory, 
  			dnn = c("Ethnicity(2)", "Ethnicity(5)"))
  
```

###Nationality###

```{r, echo = FALSE, warning = FALSE}


  table(data$Nationality, dnn = "Nationality")  

  table(data$CountryofOrigin, dnn = "Country of Origin")

  mapColor <- c("#E0A1AC", "#E0A1AC", "#E0A1AC", "#E0A1AC", "#E0A1AC",
  							"#E24444", "#E31C17")

  mapCountryData(map, nameColumnToPlot= "MaxCount", catMethod = "categorical",
  							 colourPalette = mapColor, mapTitle = "", addLegend = F,
  							 missingCountryCol = "#bdbdbd")

```

```{r, echo = FALSE, warning = FALSE}

 table(data$OccupationClassification, dnn = "Social Class (mothers)")

```


```{r, echo = FALSE, warning = FALSE}

 table(data$OccupationClassification.1, dnn = "Social Class (partner)")

```

```{r, echo = FALSE, warning = FALSE}

 table(data$PatientType, dnn = "Patient type")

```

```{r, echo = FALSE, warning = FALSE}

   table(data$MaritalStatus, data$partner, 
   			 dnn = c("Marital status", "Any partner"))

```

Folic acid 

```{r, echo = FALSE, warning = FALSE}


	table(data$Folicacidintake, data$Appropriatedose, 
				dnn = c("Took folic acid", "Appropriate dose"))

```

Multivitamin

```{r, echo = FALSE, warning = FALSE}

  table(data$Multivitamin, dnn = "Multivitamin")

```

Smoking

```{r, echo = FALSE, warning = FALSE}

  table(data$Smoking, data$NoofCiggarettesperday,
  			dnn = c("Smoking", "Number per day"), useNA = "always")
  			

```

Drinking

```{r, echo = FALSE, warning = FALSE}

  table(data$Alcohol, data$Noofunitsweek, 
  			dnn = c("Alcohol", "Units per week"))

```

Previous obsteric history

```{r, echo = FALSE, warning = FALSE}

  table(data$Previousrelaventobstetrichistory)
  table(data$Previousrelaventobstetrichistory.1)

```


```{r, warning = FALSE}

  	qplot(factor(data$Parity)) # Looks good

# Gravida ###

	table(data$Gravida)
	qplot(factor(data$Gravida))

	table(data$Parity, data$Gravida)

# 4 observations in lower diagonal, where they should not be.

# Fetal loss ####
# Not sure if this is the best way to do this)

	data$fetalLoss <- data$Gravida - data$Parity - 1

	qplot(factor(data$fetalLoss))
	table(factor(data$fetalLoss))

# 4 with negative values - each of these 4 have a previous history of multiple
# pregnancy (Historyofmultiplepregnancy = 1 (is that what this means?))

	data[data$fetalLoss == -1, ]

	table(data$Historyofmultiplepregnancy) # Not sure on coding here 0, 1, 2.

	data[data$fetalLoss == 0 & data$Historyofmultiplepregnancy == 1, ]

# 2 people, IDs 33 and 358   - Should these be coded to reflect a prior fetal loss TODO


# Birth outcomes ##############################################################
# Gestational age ####

	table(data$Gestationatdelivery) # One with a zero TODO

	table(data$Gestationalageatdeliveryday) # Why is this ever > 6? TODO

	data$gestAge <- data$Gestationatdelivery + (data$Gestationalageatdeliveryday / 7)

	summary(data$gestAge)
	qplot(data$gestAge)

# Normal vaginal delivery ####

  table(data$NVD)

# Induction ####

  table(data$Induction)

  table(data$NVD, data$Induction)

# Operative delivery ####

  table(data$Operativedelivery)

  table(data$Operativedelivery, data$NVD) # 2 classified as NVD and Operative TODO

  data[data$Operativedelivery >0 & data$NVD == 1, ]

# Live Births ####

  table(data$LiveInfant)


```


